---
title: "condFDR_analysis"
author: "Gokberk Alagoz"
date: "November 11, 2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(condFDR)
library(dplyr)
library(tidyr)
library(ggplot2)
library(qqman)

```

### Number of overlapping SNPs in dyslexia and clap-beat summary statistics

There are 8,288,850 SNPs in the clap-beat, 7,995,922 SNPs in the dyslexia summary statistics files. *7,027,342* of these SNPs are present in both files.

```{r echo = FALSE, warning=FALSE}
clapbeat = fread("/data/clusterfs/lag/users/gokala/beat-dyslexiaevol/data/clap-beat_overlap_forManhattan4.tab")
dyslexia = fread("/data/clusterfs/lag/users/gokala/beat-dyslexiaevol/data/dyslexia_overlap_forManhattan4.tab")
colnames(clapbeat) = c("SNP", "CHR", "BP","clapbeat_p")
colnames(dyslexia) = c("SNP", "CHR", "BP","dyslexia_p")

for_clapbeat_cFDR = clapbeat
for_dyslexia_cFDR = dyslexia
for_clapbeat_cFDR$dyslexia_p = dyslexia$dyslexia_p
for_dyslexia_cFDR$clapbeat_p = clapbeat$clapbeat_p

nrow(clapbeat[clapbeat$clapbeat_p < 1E-5,])
nrow(dyslexia[dyslexia$dyslexia_p < 1E-5,])

```

### Conditional and conjunctional FDR analysis with the _condFDR_ package

```{r condFDR analysis, echo = FALSE, warning=FALSE}

## Clap-beat GWAS
# estimate the conditional FDR for p1 (clapbeat), conditioned on p2 (dyslexia)

clapbeat_cFDR = cFDR(for_clapbeat_cFDR, "clapbeat_p", "dyslexia_p", p2_threshold = 1E-5)

# estimate the conjuctional conditional FDR for p1 and p2, 
# corresponding to a false discovery rate for SNPs that 
# are associated with both phenotypes

clapbeat_ccFDR = ccFDR(for_clapbeat_cFDR, "clapbeat_p", "dyslexia_p", p_threshold = 1E-5)
clapbeat_ccFDR_nothreshold = ccFDR(for_clapbeat_cFDR, "clapbeat_p", "dyslexia_p")

## Dyslexia GWAS
# estimate the conditional FDR for p1 (dyslexia), conditioned on p2 (clapbeat)

dyslexia_cFDR = cFDR(for_dyslexia_cFDR, "dyslexia_p", "clapbeat_p", p2_threshold = 1E-5)

# estimate the conjuctional conditional FDR for p1 and p2, 
# corresponding to a false discovery rate for SNPs that 
# are associated with both phenotypes

dyslexia_ccFDR = ccFDR(for_dyslexia_cFDR, "dyslexia_p", "clapbeat_p", p_threshold = 1E-5)
dyslexia_ccFDR_nothreshold = ccFDR(for_dyslexia_cFDR, "dyslexia_p", "clapbeat_p")

hist(-log10(clapbeat_ccFDR$ccFDR), col = "green", main = "-log10(pval) histograms")
hist(-log10(dyslexia_ccFDR$ccFDR), col = "red", add = T)

hist(-log10(clapbeat_ccFDR$cFDR1), col = "green", main = "-log10(pval) histograms")
hist(-log10(dyslexia_ccFDR$cFDR1), col = "red", add = T)

hist(-log10(clapbeat_ccFDR$cFDR2), col = "green", main = "-log10(pval) histograms")
hist(-log10(dyslexia_ccFDR$cFDR2), col = "red", add = T)


```
Here I followed the default p-value threholds provided in the Github tutorial, but they mention that these thresholds are rather arbitrary, so we can change these for future analysis.

#### condFDR
A p-value threshold of 1e-5 was applied on the p-values that we condition on. This yields `r nrow(clapbeat_cFDR)` SNPs when we condition clap-beat p-values on dyxlexia p-values, and `r nrow(dyslexia_cFDR)` SNPs when dyslexia is conditioned on clap-beat. Here I do not apply a threshold on conditional FDRs.

#### conjFDR
Here I applied the same threshold (1e-5) on conjunctional FDR p-values and ended up with `r nrow(clapbeat_ccFDR)` SNPs for clap-beat and `r nrow(dyslexia_ccFDR)` SNPs for dyslexia with a conjunctional FDR value less than 1e-5.

### Clap-beat GWAS, conditional and conjunctional FDR of clap-beat on dyslexia

```{r echo = FALSE, fig.width=8, fig.height=8, fig.align = "center", warning=FALSE}

# raw clapbeat manhattan plot
clapbeat_manhattan = clapbeat[, c("SNP", "CHR", "BP", "clapbeat_p")]
colnames(clapbeat_manhattan)[4] = "P"

clapbeat_manhattan = clapbeat_manhattan[clapbeat_manhattan$CHR %in% seq(1:22), ]
clapbeat_manhattan$CHR = as.numeric(clapbeat_manhattan$CHR)
clapbeat_manhattan$BP = as.numeric(clapbeat_manhattan$BP)


# condFDR clapbeat manhattan plot
clapbeat_cFDR_manhattan = clapbeat_cFDR[, c("SNP", "CHR", "BP", "cFDR")]
colnames(clapbeat_cFDR_manhattan)[4] = "P"

clapbeat_cFDR_manhattan = clapbeat_cFDR_manhattan[clapbeat_cFDR_manhattan$CHR %in% seq(1:22), ]
clapbeat_cFDR_manhattan$CHR = as.numeric(clapbeat_cFDR_manhattan$CHR)
clapbeat_cFDR_manhattan$BP = as.numeric(clapbeat_cFDR_manhattan$BP)

# conjFDR clapbeat manhattan plot
clapbeat_ccFDR_manhattan = clapbeat_ccFDR[, c("SNP", "CHR", "BP", "ccFDR")]
colnames(clapbeat_ccFDR_manhattan)[4] = "P"

clapbeat_ccFDR_manhattan = clapbeat_ccFDR_manhattan[clapbeat_ccFDR_manhattan$CHR %in% seq(1:22), ]
clapbeat_ccFDR_manhattan$CHR = as.numeric(clapbeat_ccFDR_manhattan$CHR)
clapbeat_ccFDR_manhattan$BP = as.numeric(clapbeat_ccFDR_manhattan$BP)

# plot

par(mfrow=c(3,1))
manhattan(clapbeat_manhattan, annotatePval = 0.00001, main = "Clap-beat GWAS",
          col = c("blue4", "orange3"), chrlabs = c(1:22))
manhattan(clapbeat_cFDR_manhattan, annotatePval = 0.00001, main = "Conditional FDR on dyslexia",
          col = c("blue4", "orange3"), chrlabs = c(1:22))
manhattan(clapbeat_ccFDR_manhattan, annotatePval = 0.00001, main = "Conjunctional FDR on dyslexia",
          col = c("blue4", "orange3"), chrlabs = c(1:22))

```

### Dyslexia GWAS, conditional and conjunctional FDR of dyslexia on clap-beat

```{r echo = FALSE, fig.width=8, fig.height=8, fig.align = "center", warning=FALSE}

# raw dyslexia manhattan plot
dyslexia_manhattan = dyslexia[, c("SNP", "CHR", "BP", "dyslexia_p")]
colnames(dyslexia_manhattan)[4] = "P"

dyslexia_manhattan = dyslexia_manhattan[dyslexia_manhattan$CHR %in% seq(1:22), ]
dyslexia_manhattan$CHR = as.numeric(dyslexia_manhattan$CHR)
dyslexia_manhattan$BP = as.numeric(dyslexia_manhattan$BP)

# condFDR dyslexia manhattan plot
dyslexia_cFDR_manhattan = dyslexia_cFDR[, c("SNP", "CHR", "BP", "cFDR")]
colnames(dyslexia_cFDR_manhattan)[4] = "P"

dyslexia_cFDR_manhattan = dyslexia_cFDR_manhattan[dyslexia_cFDR_manhattan$CHR %in% seq(1:22), ]
dyslexia_cFDR_manhattan$CHR = as.numeric(dyslexia_cFDR_manhattan$CHR)
dyslexia_cFDR_manhattan$BP = as.numeric(dyslexia_cFDR_manhattan$BP)

# conjFDR dyslexia manhattan plot
dyslexia_ccFDR_manhattan = dyslexia_ccFDR[, c("SNP", "CHR", "BP", "ccFDR")]
colnames(dyslexia_ccFDR_manhattan)[4] = "P"

dyslexia_ccFDR_manhattan = dyslexia_ccFDR_manhattan[dyslexia_ccFDR_manhattan$CHR %in% seq(1:22), ]
dyslexia_ccFDR_manhattan$CHR = as.numeric(dyslexia_ccFDR_manhattan$CHR)
dyslexia_ccFDR_manhattan$BP = as.numeric(dyslexia_ccFDR_manhattan$BP)

# plot

par(mfrow=c(3,1))
manhattan(dyslexia_manhattan, annotatePval = 0.00001, main = "Dyslexia GWAS",
          col = c("blue4", "orange3"), chrlabs = c(1:22))
manhattan(dyslexia_cFDR_manhattan, annotatePval = 0.00001, main = "Conditional FDR on clap-beat",
          col = c("blue4", "orange3"), chrlabs = c(1:22))
manhattan(dyslexia_ccFDR_manhattan, annotatePval = 0.00001, main = "Conjunctional FDR on clap-beat",
          col = c("blue4", "orange3"), chrlabs = c(1:22))

```

```{r}
sessionInfo()
```

